---
title: "lecture abd lab 8"
output: pdf_document
date: "2023-03-07"
---
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
```

```{r}
dgp <- function (n) tibble(Z = runif(n)) %>% 
  summarize(mean = mean(Z))

density <- function(d) {
  d %>% ggplot() +
    geom_density(aes(x=mean))
}

sample_means <- cbind(n_10=map_df(1:100,\(x)dgp(10)),
                       n_50=map_df(1:100,\(x)dgp(50)),
                       n_100=map_df(1:100,\(x)dgp(100)),
                       n_150=map_df(1:100,\(x)dgp(150)),
                       n_250=map_df(1:100,\(x)dgp(250))
)
sample_means


n10 <- map_df(1:100,\(x)dgp(10))
n50=map_df(1:100,\(x)dgp(50))
n100=map_df(1:100,\(x)dgp(100))
n150=map_df(1:100,\(x)dgp(150))
n250=map_df(1:100,\(x)dgp(250))
               
density(n10)
density(n50)
density(n100)
density(n150)
density(n250)

rnorm(100) %>% 
  sample(10,replace = TRUE)

```

bootstrapping
```{r}
boot_dif <- function(df, s) {
s_mean <- mean(df) 
strap_mean <-  df %>% 
    sample(size = s, replace = TRUE) %>% 
  mean()
return(error <- s_mean-strap_mean)
}

boot <- function(n,df,s) {
 error <-  map_dbl(1:n,\(x)boot_dif(df,s))
 return(tibble(
   sample_mean = mean(df),
   upper_bound = sample_mean - quantile(error,probs=.025),
   lower_bound = sample_mean - quantile(error,probs=.975)
 ))
}

test <- tibble(X=rnorm(100))
yup <- boot(10000, test$X ,100)
yup

```

# LAb 8

SE derivation for known Variance: $\sigma^2$
$\psi = \bar{X}$
$\bar{X} = \frac{1}{n} \sum X_i$
$var(\bar{X}) = var(\frac{1}{n}\sum X_i)$
$var(\bar{X}) = \frac{1}{n^2}var(\sum X_i) $
$var(\bar{X}) = \frac{1}{n^2}\sum var(X_i) $
$var(\bar{X}) = \frac{1}{n^2} * n*\sigma^2 $
$var(\bar{X}) = \frac{n\sigma^2 }{n^2}  $
$var(\bar{X}) = \frac{\sigma^2 }{n}$
$SE(\bar{X}) = \frac{\sigma}{\sqrt{n}}$

CLT: if IID we good estimator is asymptotically normal

Delta method: 

for transformations of estimate to get transformed sampling distribution use delta method:

$g(\hat{X}) = N(g(\mu),\sigma^2*(g'(\mu))^2)$


$ \bar{X} \frac{+}{-} 1.96\frac{\sigma}{\sqrt{n}}$

Bootstrapping
no want to derive SE formula or can't
1. Take sample calc esitmate
2. resample from sample with replacement, calc estimate
3. repeat
4. have a sampling distribution from list of estimates
5. take variance of that and take square root to get SE.


```{r}
estimator = function(data){ var(data$X)}
bootstrap_estimator = function(data, estimator, B=100) {
estimate = estimator(data)
samples <- map_df(1:B,(.)function{slice_sample(data,prop =1,replace = T)})  # generate B bootstrap samples, estimate, compute quantiles...
tibble(
estimate = estimate, # the point estimate
ci_a, # lower bound of the CI
ci_b # upper bound of the CI
)
}
```

```{r}
testy <- tibble(X= rnorm(1000))
samples <- map_dbl(1:10,function(.){slice_sample(testy,prop =1,replace = T) %>% 
    estimator()})
```

