---
title: "Hw4 - Beimnet Taye"
output: pdf_document
date: "2023-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rmutil)
library(pracma)
```
# Worked with Joan Shim and Lucas Yoshida
## P1

### 1

-   They are identical, identically distributed, not independent.

### 2

-   B: They are identically distributed and independent. C: They are identically distributed and independent.

### 3

-   A and B

### 4

-   all three have the same marginal distribution for X

## P2

### 1

-   Random Variable: Let $C$ be a random variable that takes an individual and assess whether they are infected with covid or not.

-   Domain: $f^{-1}_C = \{individual\}$

-   Range: $F_C(individual) = \{I = infected, H = Healthy\}$

-   Random Variable: Let $t$ be a random variable that takes an individual and assess whether they have tested positive or negative.

-   Domain: $f^{-1}_t = \{individual\}$

-   Range: $F_t(individual) = \{+ = positive, - = negative \}$

-   $P(t =+) = 0.02 \text{, } P(C = I\cap t = +) = 0.01$

-   $P(t =+)$ is a marginal probability and $P(C = I\cap t = +) = 0.01$ is a joint probability

```{r}
pp <- 0.02
ppi <- 0.01
```

### 2

-   $P(C = I | t = +) = \frac{P(C = I, t =+)}{P(t = +)}$

```{r}
pigp <- ppi/pp
pigp
```

### 3

-   $P(t = +|C=I) = 0.9$
-   $P(C = I) = 0.009$

```{r}
ppgi <- 0.9
pi <- 0.009
```

### 4

$P(I|+) = \frac{P(I) * P(+|I)}{P(+)}$

```{r}
pigp <- (pi*ppgi)/pp
pigp
```

## P3

### 1

-   They are identically distributed. Not independent.

### 2

-   They are identical, identically distributed, and not independent.

## P4

### 1

```{r}
X = function() {
  c(X = rnorm(1))
}
Y_given_X = function(x) {
  c(Y = rexp(1, rate=abs(x)))
}
Y_and_X = function() {
  x = X()
  y = Y_given_X(x)
  return(list(X=x, Y=y))
}
sample = function(dist, n, ...) {
  map_df(1:n, function(i) dist(...))
}

prob_x <- X %>% 
  sample(10000) %>% 
  count(X<0) %>% 
  mutate(p = n/sum(n))

pull(prob_x[2,3])

prob_ygx <- Y_given_X %>% 
  sample(10000, x= 1) %>% 
  count(Y>1) %>% 
  mutate(p = n/sum(n))

pull(prob_ygx[2,3])

prob_yandx <- Y_and_X %>% 
  sample(10000) %>% 
  count(X<0 & Y > 1) %>% 
  mutate(p = n/sum(n))

pull(prob_yandx[2,3])
```

### 2

-   It's not possible to compute with the given functions alone without manipulation since Y is continuous and we can't get it to exactly equal 1 when we sample from the joint and thus the probability using this sampling code is zero.

### 3

-   Sample from the joint function and filter for when Y > 1 to get the probability when X < 0.

```{r}
prob_xgy1 <- Y_and_X %>% 
  sample(10000) %>% 
  filter(Y > 1) %>%
  count(X<0) %>% 
 mutate(p = n/sum(n))

pull(prob_xgy1[2,3])
```

## P5

### 1

```{r}
f_X = function(x) {dnorm(x)
}

f_Y_given_X = function(y,x) {
  dexp(y, rate=abs(x))
}

f_Y_and_X <- function(y,x){
  X <-  f_X(x)
  Y <- f_Y_given_X(y,x)
  return(c(X,Y))
  
}
```

### 2

```{r}
f_Y_given_X_is_0 <- function(y){
  dexp(y, rate=abs(0))
}
```

### 3

```{r}
p1 <- integrate(f_X,-1,0)

p2 <- integrate(f_Y_given_X_is_0,-1,0)

p3 <- int2(f_Y_and_X, a = c(0,0), b = c(1,1))

p1[["value"]]
p2[["value"]]
p3
```
