---
title: "lec and lab week 6"
output: pdf_document
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
```

## LECTure

```{r}
yap <- tibble(
  w = 1:1000,
  X = case_when(
    w %% 2 == 0 ~ 0,
    w %% 3 == 0 ~ 1,
    TRUE        ~ 2
  )
)
E <- yap %>% 
  summarize(mean = mean(X))

yap_pdf <- yap %>% 
  count(X) %>% 
  mutate(p = n/nrow(yap),
         fxx = p * X ) %>% 
  summarize(E = sum(fxx))

yap %$% {
  mean(X)
}

integrand <- function(x) {
  x*dexp(x,2)
}

integrate(integrand,0,999)

```


```{r}
dgp <- function(n){
tibble(
  X = rnorm(n,-0.5,1),
  Y = runif(n),
  Z = X+Y
)
}

yeet <- dgp(10000)
d_e <- dgp(100000) %>% 
  summarize(EX = mean(X),
            EY = mean(Y),
            EZ = mean(Z)) #they are additive inverses
Density <- function(df, ...){
  df %>% 
    ggplot() +
    geom_density(aes(x=...))
}

Density(yeet, X)
Density(yeet, Y)
Density(yeet, Z)
```
```{r}
con <- dgp(10000) %>% 
  filter(Z==0) %>% 
  summarize(mean(X))

```

```{r}
DGP1 <- function(n){
  tibble(
    X = runif(n),
    Y = rnorm(n,X)
  )
}

test <- DGP1(100)
yup1 <- DGP1(10000) %>% 
  group_by(round(X, 2)) %>% 
  summarize(mean = mean(Y))
```

# LAB

Expectation as weighted average with pdf densities (relative frequency that the value x occurs) as weights to value x.

weighted ave = weight(1) * 1 + weight(2) * 2 .... weight(n) * n

discrete var expected value: pdf(x1) * x1 + pdf(x2) + x2 ... pdf(xn) + xn

continuous var expected value: integral(x*pdf(x))
  - when integrating a symmetric distribution expectation is the symmetry axis since areas will cancel out for every value x barring the symmetry axis. 
  
expectation of a function 

densities map over from any transformation x when passing through a function so E[g(x)] = g(x) * pdf(x). Can get expectation with pdf of X for any transformation function of X, g(X).

independence rules also apply to expectation:

E[X*Y] = E[X] * E[Y] if X and Y are independent 







