---
title: "lecture and lab week 3"
output: html_document
date: "2023-02-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# LECTURE

prob1

```{r}
w <- seq(0,1,length = 1000)

#A <- seq(0,0.3, length = 1000)

RV_x <- tibble("universe" = w, "X" = sin(10*w))
RV_y <- tibble("universe" = w, "Y" = sin(10*(1-w)))

prob_x <- RV_x %>% 
  mutate(inside = X <= 0.3 | X >= 0.6) %>% 
  count(inside) %>% 
  mutate(probability = n/sum(n))

prob_y <- RV_y %>% 
  mutate(inside = Y <= 0.3 | Y >= 0.6) %>% 
  count(inside) %>% 
  mutate(probability = n/sum(n))

# They output the same probabilites!

hist_x <- RV_x %>% 
  ggplot(aes(x=X)) +
  geom_histogram()

hist_y <- RV_y %>% 
  ggplot(aes(x=Y)) +
  geom_histogram()

hist_y

```

cumulative distrib function

```{r}
cumul <- function(sample, a) {
 prob <-  sample %>% 
    mutate(inside = X < a) %>% 
    count(inside) %>% 
   mutate(probs = n/sum(n))
 
 return(pull(prob[2,3]))
}

lol <- cumul(RV_x, 0.5)
RV_x_cumul <- RV_x %>% 
  arrange(X) %>% 
  mutate(Fx = row_number()/n()) %>% 
  ggplot() +
  geom_line(aes(x = X, y = Fx))

RV_x_cumul


```

## R FUNCTIONS FOR DISTRIBTUINOS

### NORMAL

```{r, eval = F}
pnorm(x, mean, sd) #CDF
dnorm()       # PDF
qnorm() #Qf
rnorm() #sampling draw x amount of samples from distribution
```

###Uniform

```{r, eval = F}
punif(x, min, max) #CDF
dunif()
qunif()
```

### BERNOULI

```{r, eval = F}
pbinom(x, 1, prob)
```

```{r}
exp <- tibble(X= seq(-10,10, length=1000),
              Fx = dexp(X, rate = 1)) %>% 
  ggplot() +
  geom_line(aes(x = X, y = Fx))
exp

exp_h <- tibble(Fx = rexp(10000, rate = 1)) %>% 
  ggplot() +
  geom_histogram(aes(x = Fx))
exp_h
```
# LAB

## 1

### 1.1

A random variable takes a universe and outputs a characteristic measurement of said universe. domain Universe. Range Number.

### 1.2

An event is a subset of all the possible universes

### 1.3

Fx -> subtract Fx(a) from Fx(b). 
fx -> calculate area under curve over interval (a,b]

### 1.4

Standard Normal Distribution


## 2

### 2.1

```{r}
pnorm(3) + pnorm(3,lower.tail=F)
```

### 2.2
```{r}
ye <- tibble(x = seq(-10,20, length = 100),
             inside = as.numeric(x>= 0 & x<= 10)) %>% 
  ggplot() +
  geom_line(aes(x = x, y = inside))
ye

```
Because its area is greater than one (PDF)

## 3
still pnorm?
```{r}
yap <- tibble(X = rnorm(10000),
              Y = X^2,
              inside = Y < 0.5) %>% 
  count(inside) %>% 
  mutate(prob = n/sum(n)) 
Y_prob <-  pull(yap[2,3])

X_prob <- pnorm(0.5)

df <- tibble(X = rnorm(10000),
              Y = X^2,
              inside = Y < 0.5)

CDF_Y <- df %>% 
  arrange(Y) %>% 
  mutate(prob = row_number()/nrow(df))

CDF_X <- df %>% 
  arrange(X) %>% 
  mutate(prob = row_number()/nrow(df))

Y_line <- CDF_Y %>% ggplot() +
  geom_line(aes(x = Y, y = prob))

X_line <- CDF_X %>% ggplot() +
  geom_line(aes(x = X, y = prob))

X_line

Y_line
  
```
## 4

```{r}
anotha <- tibble(w= (1:10),
                 X = w/2 + 1) %>% 
  arrange(X) 

anotha <- anotha %>%
  mutate(prob = row_number()/nrow(anotha))

CDF_anotha <- anotha %>% 
  ggplot() +
  geom_point(aes(x=X, y = prob))
CDF_anotha

## big boi
anotha_b <- tibble(w= seq(0,10, length = 10000),
                 X = w/2 + 1) %>% 
  arrange(X) 

anotha_b <- anotha_b %>%
  mutate(prob = row_number()/nrow(anotha_b))

CDF_anotha_b <- anotha_b %>% 
  ggplot() +
  geom_point(aes(x=X, y = prob))
CDF_anotha_b

prob_4 <- anotha_b %>% 
  filter(X > 3.9 & X <= 4)
```
## 5
```{r}

```

