---
title: "PURRR"
output: html_document
date: "2023-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# PURRRRR

# MAP FUNCTION ITERATES

map_lgl for logicals

map_int for integers

map_dbl for doubles

map_cr for characters

- these work if your function returns single values

- if your only ever going to use a function iteratively you can create an anynymous function
  - a function solely defined in the map


```{r}

my_fn <- function(x,y){ 3 + x+ y}
my_val <- 1:20
map_dbl(my_val, function(x){
  3 * x + 2
})

# VS

map_dbl(my_val, ~ 3 * .x + 2)
```

Both are the same way to define an anon function, ~ better for one liners

you can pass other arguments of function into map. good to specificy them. First arguement is the first arguemetnt of the function you are passing (.x)

```{r}
map_dbl(my_val,my_fn, y= 2)
```

With dfs. Can iterate non vectorized functons like read_csv. map_df auto binds rows by retureing a data frame.
.id allows you to define specfic source of observations

```{r}

```

