---
title: "week 1"
output: html_document
date: "2023-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sets)

```

### SETS AND FUNCTIONS

# Notation

M = {a,b,c}

b e M

order doesn matter

no duplicates, can be infinite 

N = set of natural numbers

R = all Real numbers

# parts

subset - a set of a larger set

element - item in a set


Sets can share elements
 - ven diagram
 
 - Union (U)
    - what us in either set
    
  - intersection (n)
    - in both
    
The complement is everything NOT in the set

  - A^c = everything not in set A in comparision to some larger universal set. 
  
```{r}
A <- set(2,1,8)
B <- set(9,2,6)

#comp_union = function(A,B,universe = (0:9)) {}

```
  
  
  
(A U B)^c = A^c n B^C


C = {a, p, f} C < A U P U F

ec e! C

## Functions

relationship between elements in one set with elements in another

a box of algorithims 

every input must get exactly one output

set of inputs: Domain

set of outputs: Range

f: D -> R

Domain and range can be the same real number -> real number


If Domain and range are ordered you can draw a graph (D,R)

functions in R  


# Set indicator functions 

  - gives true or false if an inputted element is in a given set D: elements R: True/false
  - graph block graph range of 0 and 1 
  
# multiple arguements 
 - two inputs one output
 - graph: a 2d surface
 - ex. indicator set of coordinates in a plane (x and y)
 
 - intersection as a two arg function


### LAB

```{r}
position <- function(t){
  
   y <- 5*cos(t * 2 * pi / 12)
   
   return(y)
  
  
}


position(3)
```


```{r}
position_all <- function(t,l,r) {
  
  
  hr <- 12
  min <- 720
  sec <- 43200
  
  if(r == "hr") {
      y <- l*cos(t * 2 * pi / 12)
  } else if( r == "min") {
    y <- l*cos(t * 2 * pi / 720)
  } else (r == "sec") {
    
     y <- l*cos(t * 2 * pi / 43200)
  }
  
  
  
   
   return(y)
  
  
  
}

position_all(0,5,hr)


```
Function factories
```{r}


f_factory <- function(l=5,r=12){
  
  function(t){
    l*cos(t * 2 * pi / r)
  }
}

f_factory(3,12)(6)  #f_factory calls a function in which the second parenthetical is passed.

```


 git config --global user.email "btaye@berkeley.edu"

